Accelerating Deep Neural Network Training on FPGAs Facilitating the Use of Variable Precision

### 2018-07-01 SWoPP2018 Paper/paper.tex
Added CIFAR training and inference.
Accuracy is really bad.

\subsection{Network Tested + Data-set}
The used example network is a simple multilayer perceptron.
The main dataset used was MNIST, so 784 input features.
The structure of the sample MNIST network is.
\begin{enumerate}
  \item Dense(512)
  \item Relu
  \item Dense(512)  
  \item Relu
  \item Dense(10)
  \item Softmax
\end{enumerate}
So in total a three layer network, or threes sets of weight matrices and bias vectors.
The data set was trained using the first \num{50000} images from the MNIST dataset.
The last \num{10000} images are used as the test data set.
Training was done using a batch size of \num{128}.
The CPU code was implemented using the NumPy array.dot function.

\section{Results}\label{sec:result}
The host (numpy based) code benchmarked at \num{60} GFLOPS for simple inference.
A first naive implementation clocked about \num{2.5} GFLOPS for this problem size.
The results for the different final FPGA kernels are shown in \cref{tab:kernel-performance}.

\begin{table}[h]
  \centering
  \caption{FPGA Kernel performance}\label{tab:kernel-performance}
  \begin{tabular}{ll}
    \toprule
    \textbf{Block size} & \textbf{GFLOPS}  \\
    \midrule
    128$\times$128 & \num{0}  \\ % Too big
    100$\times$100 & \num{0} \\ % Too big (86 is also too big)
    84$\times$84 & \num{0}  \\
    64$\times$64 & \num{17.8}  \\ % kernel_nn_blocking_B64inorder_hw_2018-06-28T22.52.27.aocx Network: [28 * 28, 512, 512, 10]
    32$\times$32 & \num{0}  \\
    16$\times$16 & \num{4.3}  \\ % kernel_nn_blocking_B16inorder_hw_2018-06-28T22.53.27.aocx Network: [28 * 28, 512, 512, 10]
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[h]
  \centering
  \caption{FPGA Hardware utilisation for different blocksizes}\label{tab:fpga-util}
  \begin{tabular}{lllll}
    \toprule
    \textbf{Block size} & \textbf{Logic} & \textbf{DSPs} & \textbf{RAMs} & \textbf{Fmax} \\
    \midrule
    128$\times$128 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    100$\times$100 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    84$\times$84 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    64$\times$64 & \SI{33}{\percent} & \SI{10}{\percent} & \SI{68}{\percent} & \SI{172.8}{\mega\hertz}  \\ % kernel_nn_blocking_B64inorder_hw_2018-06-28T22.52.27.aocx
    32$\times$32 & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\percent} & \SI{0}{\mega\hertz}  \\
    16$\times$16 & \SI{17}{\percent} & \SI{3}{\percent} & \SI{30}{\percent} & \SI{209.16}{\mega\hertz}  \\ % kernel_nn_blocking_B16inorder_hw_2018-06-28T22.53.27.aocx
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Training Performance}
The accuracy after training on the CPU using Keras is \SI{97.9}{\percent}.
The FPGA gets slightly different floaing point results due to rounding differences but is matches the accuracy.

\section{Conclusion}\label{sec:conclusion}

>>>>

\subsection{Network Tested + Data-set}
The used example network is a simple multilayer perceptron.
The first dataset used was MNIST, so \num{784} input features.
The structure of the sample MNIST network is.
\begin{enumerate}
  \item Dense(512)
  \item Relu
  \item Dense(512)  
  \item Relu
  \item Dense(10)
  \item Softmax
\end{enumerate}
So in total a three layer network, or threes sets of weight matrices and bias vectors.
The data set was trained using the first \num{60000} images from the MNIST dataset.
The last \num{10000} images are used as the test data set.
The second dataset used was CIFAR10, so \num{3072} input features.
The structure of the sample CIFAR10 network is.
\begin{enumerate}
  \item Dense(1024)
  \item Relu
  \item Dense(512)  
  \item Relu
  \item Dense(512)  
  \item Relu
  \item Dense(10)
  \item Softmax
\end{enumerate}

The data set was trained using the first \num{50000} images form the dataset.
The last \num{10000} images were used as the test data set.
All training was done using a batch size of \num{128}.
The CPU code was implemented using the NumPy array.dot function.

\section{Results}\label{sec:result}
The host (numpy based) code benchmarked at \num{60} GFLOPS for simple inference of the first network and about \num{160} GFLOPS for the second network.
A first naive implementation on the FPGA clocked about \num{2.5} GFLOPS for this problem size.
The results for the different final FPGA kernels are shown in \cref{tab:kernel-performance}.
Larger block sizes would exhaust the block RAM on the used FPGA device.
The next logical size up from \num{64} would be \num{84} (\num{3}$\cdot$\num{28}), this would use to much of the block RAM already.

\begin{table}[h]
  \centering
  \caption{FPGA Kernel performance}\label{tab:kernel-performance}
  \begin{tabular}{llll}
    \toprule
    \textbf{Block size} & \textbf{GFLOPS MNIST}  & \textbf{GFLOPS CIFAR10}  \\
    \midrule
    64$\times$64 & \num{17.8} & \num{18.9}  \\ % kernel_nn_blocking_B64inorder_hw_2018-06-28T22.52.27.aocx Network MNIST: [28 * 28, 512, 512, 10] CIFAR: [32 * 32 * 3, 1024, 512, 512, 10]
    16$\times$16 & \num{4.3} & \num{4.4} \\ % kernel_nn_blocking_B16inorder_hw_2018-06-28T22.53.27.aocx Network MNIST: [28 * 28, 512, 512, 10] CIFAR: [32 * 32 * 3, 1024, 512, 512, 10]
    \bottomrule
  \end{tabular}
\end{table}

The performance is memory bandwidth limited for these kernels.

@@

\begin{table}[h]
  \centering
  \caption{FPGA Hardware utilisation for different blocksizes}\label{tab:fpga-util}
  \begin{tabular}{lllll}
    \toprule
    \textbf{Block size} & \textbf{Logic} & \textbf{DSPs} & \textbf{RAMs} & \textbf{Fmax} \\
    \midrule    
    64$\times$64 & \SI{33}{\percent} & \SI{10}{\percent} & \SI{68}{\percent} & \SI{172.8}{\mega\hertz}  \\ % kernel_nn_blocking_B64inorder_hw_2018-06-28T22.52.27.aocx
    16$\times$16 & \SI{17}{\percent} & \SI{3}{\percent} & \SI{30}{\percent} & \SI{209.16}{\mega\hertz}  \\ % kernel_nn_blocking_B16inorder_hw_2018-06-28T22.53.27.aocx
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Training Performance}
The MNIST network accuracy after training on the CPU using Keras is \SI{97.9}{\percent}.
And the CIFAR10 accuracy after training is \SI{47.9}{\percent}
The FPGA gets slightly different floaing point results due to rounding differences but is matches the accuracy.

\section{Conclusion}\label{sec:conclusion}
This framework makes for a good basis to start some research in different number
